News Across Languages - Cross-Lingual Document Similarity and Event Tracking

### 2015-03-24 rupnik2015.tex
typos

Currently we compute similarities over $24$ languages with tags: eng, spa, deu, zho, ita, fra, rus, swe, nld, tur, jpn, por, ara, fin, ron, kor, hrv, tam, hun, slv, pol, srp, cat, ukr but we support up to $100$ top Wikipedis languages. Our data stream is Newsfeed (http://newsfeed.ijs.si/) that provides $430k$ unique articles per day. Our system currently computes 400k similiarites per 200ms, that means that we compute $16 \cdot 10^{10}$ similarities per day. We
store one day buffer for every language, for that we need 1.5 GB of memory, we store documents as 500 dimensional vectors. We must note that time complexity of similarity computations scales linearly with dimension of the feature space and does not  depend of number of languages. For each article we compute top $10$  most similar ones in every other language.
For all linear algebra matrix and vector operations we use high performance numerical linear algebra libraries as BLAS, OPENBLAS and Intel MKL. This currently allows us to process more than million articles per day.
In our current implementation we use variation of hub approach. Our projector matrices are of size $500\times 300000,$ so every projector takes about $1.1$ GB of RAM. Moreover we need proxy matrices of size $500\times500$ for every language pair. That is 0.5 GB for $24$ languages and $9.2$ GB for $100$ languages. All together we need around 135 GB of RAM for the system with 100 languages.
 Usage of proxy matrices enables projection of all input documents in the common space and handling language pairs with missing or low alignment. That enables us to do block-wise similarity computations and
additional improvement in our system efficiency. Our code can be therefore easily parallelized using matrix multiplication instead of doing more matrix - vector multiplications. This speeds up our code by factor around $4.$ In this way we obtain some caching gains and ability to use vectorization.
Our system is also easily extendable. If we want to add new language we only need to compute a projector matrix and proxy matrices with all other already available languages.
Add the conclusion of the story and the wrap up, Andrej needs help with this. We presented the system ...

>>>>

Currently we compute similarities over $24$ languages with tags: eng, spa, deu, zho, ita, fra, rus, swe, nld, tur, jpn, por, ara, fin, ron, kor, hrv, tam, hun, slv, pol, srp, cat, ukr but we support up to $100$ top Wikipedia languages. Our data stream is Newsfeed (http://newsfeed.ijs.si/) that provides $430k$ unique articles per day. Our system currently computes 400k similiarites per 200ms, that means that we compute $16 \cdot 10^{10}$ similarities per day. We
store one day buffer for every language, for that we need 1.5 GB of memory, we store documents as 500 dimensional vectors. We must note that time complexity of similarity computations scales linearly with dimension of the feature space and does not  depend of number of languages. For each article we compute top $10$  most similar ones in every other language.
For all linear algebra matrix and vector operations we use high performance numerical linear algebra libraries as BLAS, OPENBLAS and Intel MKL. This currently allows us to process more than million articles per day.
In our current implementation we use variation of hub approach. Our projector matrices are of size $500\times 300000,$ so every projector takes about $1.1$ GB of RAM. Moreover we need proxy matrices of size $500\times500$ for every language pair. That is 0.5 GB for $24$ languages and $9.2$ GB for $100$ languages. All together we need around 135 GB of RAM for the system with 100 languages.
 Usage of proxy matrices enables projection of all input documents in the common space and handling language pairs with missing or low alignment. That enables us to do block-wise similarity computations and
additional improvement in our system efficiency. Our code can be therefore easily parallelized using matrix multiplication instead of doing more matrix - vector multiplications. This speeds up our code by factor around $4.$ In this way we obtain some caching gains and ability to use vectorization.
Our system is also easily extendable. If we want to add new language we only need to compute a projector matrix and proxy matrices with all other already available languages.
Add the conclusion of the story and the wrap up, Andrej needs help with this. We presented the system ...
